---
title: "Peer-graded Assignment: R Markdown and Leaflet"
author: "Patricia Londono"
date: "17/11/2020"
output: md_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Instructions
Create a web page using R Markdown that features a map created with Leaflet.

Host your webpage on either GitHub Pages, RPubs, or NeoCities.

Your webpage must contain the date that you created the document, and it must contain a map created with Leaflet. We would love to see you show off your creativity!



# Loading required packages
```{r}
library(leaflet)
library(htmltools)
```

# The World's Top Creative Cities
Data extracted from: https://www.worldatlas.com/articles/the-world-s-most-creative-cities.html 


# Reading the data and the creating the map with appropiate markers
```{r}
map <- read.csv("leaflet_data.csv",sep=",")

mymap <- map %>%
  leaflet() %>%
  addTiles() %>%
  addMarkers(popup=paste
             ("<br>Country: ", 
               htmlEscape(map$Country), 
              "<br>City: ", 
               htmlEscape(map$City), 
              "<br>Ranking: ",
               htmlEscape(map$Rank)
              ) 
            )
mymap
```



```{r}

```
